\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-13 trunk/arch.tex
updates

\rx provides a formal framework which details partitioning in
\cite{py10} and \cite{rajan12} and is not the focus of this
chapter. Coupled interaction between planning and the world on the
other hand, has a significant implication on how planning can be
integrated in such frameworks; here we show using the \eu planner can
be coupled to our framework. First we link concepts in \eu planning
with the \rx framework. Note that in the following treatise, we call
the formal semantic framework of \rx as the 'agent'. Reactors
constitute an entity within this agent and are standalone blocks of
inference.

@@

Given the above example, the \rx architecture therefore helps abstract
reactor interaction by containing them to state variables. Each
reactor is therefore agnostic where its {\em external} state
information is managed as long as this information is available and
properly maintained for managing {\em internal} state
information. This {\em internal} state in turn may be used by other
reactors. Such a formalism also constrains reactor interaction to
state information exchange which comes in two forms:

@@

\item {\em Observation on the current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so as they
  could compute their {\em internal} state variable values for this
  same tick. By doing so we propagate a consistent view of that state
  of the world throughout all the reactors as time
  advances. \comment{could do with a figure here}

\item {\em Goals request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy own
  {\em internal} objectives, a reactor may have a plan that relies on
  future values of one of its {\em external} state. When such a plan
  is identified the agent ensures that this part of the plan is given
  as a goal to the owner of the given {\em external} state
  variable(s). \comment{could do with a figure here}

@@

Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple ticks
for a reactor to produce a plan that is both complete and valid. when
such plan is identified its {\em external} part is then dispatched to
the corresponding reactors which in turn can start their deliberation
\texttt{step}s in order to satisfy these new objectives. As a
consequence physical execution will occur as these goals and planning
propagate down in the reactor graph until they eventually become goals
for the lowest level reactor which -- instead of deliberating -- may
just transform this goal in a physical command execution.

The way the agent interleave planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization happens at every single tick
and will interrupt planning allowing this reactor to identify its
current state that can propagate through the reactor graph in order to
ensure t aconsitent view of the state of the world between all the
reactors.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

This view of the reactor has two parallel tasks allow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

\subsection{The Europa deliberative reactor}
\label{sec:arch:europa}

The \eu reactor is our implementation inside \rx of a reactor that
usess automated planning for its execution. This reactor is fully
based on the europa framework and applies the core idea of \rx we
described above to implement a reactor for which the sole purpose is
to deliberate as time advance. The design of this reactor takes cues
from the IDEA architecture presented on \cite{mus02, mus06} in the
sense that both planning and execution are based on a single unified
model. The choice of \eu as the planning framework was directed by the
fact that this framework provide a rich and highly configurable
toolset to be able to infer around a rich temporal domain.

This reactor is implemented following the idea that in \rx, the
whole execution cycle is two deliberation processes. Both of these
processes are implemented as two europa solvers modifying the same
plan database and for which the execution is managed by the \rx 
framework as time advance during the execution of the system:

@@

\item The synchronization solver is a specialized europa solver that
  will integrate in the plan new information about the evolution of
  the External state variable and ensure that the reactor propagate
  these in order to both identify its current state and inform
  other reactors of any state change on its internal timelines. This
  solver is summoned at the beginning of every single tick.

\item The deliberation solver is managing the deliberation process of
  the reactor in order to produce a new plan or alter its current plan
  as new goals are given to the reactor or the synchronization solver
  identifying a conflict between the current state of the world and
  the expectations of the previous plan. This process can  span other
  multiple tick and therefore can be interrupted at any single tick in
  order for the synchronization solver to do its task.

@@

While the two processes are separate, they share the same plan
internal to a reactor and therefore at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the planning process eventually finds a
solution, synchronization is informed about generating the planned
state values on the {\em external} state variables. This {\em
  external} plan defines the set of {\em goals} for this state
variable managed by another {\em reactor}.

An important challenge for embedding automayted planning i na situated
agent as always been to go around the asumptions made by most classical
planners. Specifically one assumption restrict the problem to ``offline
planning''  defined in \cite{ghallab04} as follow:

\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{In this book $\Sigma$ identifies the world modeled
    by the plan domain} while it is planning; it plans for the given
  initial and goal states regardless of the current dynamics, if any.
\end{quotation}

This assumption, while helpfull to reduce the scope of the planning
problem, becomes problematic if one wants to embed a planner in a
real-time system that evolves in a highly dynamic environment. In a
typical robot this implies that the only point where the planner can
be informed about the current world state is right before the planning
is started by giving to the planner the intial state and current
objectives. During the planning, it is assumed that the agent can
maintin the world as perceived by the planner as stable. In
\cite{lemai04, lemai-chenevier2004}, they do attempt to reduce the
impact of planning by allowing local plan repairs which attempt to
insert repairing actions in the exisiting plan andf can be done fast
enough. Still in ythe eventuality that such repair is not possible
they add to stop their vehicle until the full (re-)planning was
complete. {\em I may need more examples/refs here}

Looking back at Fig. \ref{fig:tick-exec}, one can already see that
this asumtpion is problematic in our architecture. Indeed
synchronization that tracks state evolution can occur several times
during the reactor deliberation process. By making synchronization and
deliberation share the same plan we provide a solution that relax this
assumption. In the following sections we present how synchronization
and deliberation are implemented independently first to finally show
how their interaction through the plan helps relax this assumption by
making the planning fully integrated to the reactor execution cycle
while potentially allowing for more informed planning.

>>>>

Partitioned inference is a major contribution of \rx. However that is
not the focus of this chapter; details are in \cite{py10} and
\cite{rajan12}. Coupled interaction between planning and the world on
the other hand, has a significant implication on how planning can be
integrated in such frameworks; here we show using the \eu planner can
be coupled to our framework. First we link concepts in \eu planning
with the \rx framework. Note that in the following treatise, we call
the formal semantic framework of \rx as the 'agent'. Reactors
constitute an entity within this agent and are standalone blocks of
inference.

@@

Given the above example, the \rx architecture therefore helps abstract
reactor interaction by constraining them to state variables. Each
reactor is therefore agnostic where its {\em external} state
information is managed as long as this information is available and
properly maintained for managing {\em internal} state
information. This {\em internal} state in turn may be used by other
reactors. Such a formalism also constrains reactor interaction to
state information exchange which comes in two forms:

@@

\item {\em Observation on current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so as they
  could compute their {\em internal} state variable values for this
  same tick. By doing so we propagate a consistent view of that state
  of the world throughout all the reactors as time
  advances. \comment{could do with a figure here}

\item {\em Goal request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy own
  {\em internal} objectives, a reactor may have a plan that relies on
  future values of one of its {\em external} state. When such a plan
  is identified the agent ensures that this part of the plan is given
  as a goal to the owner of the given {\em external} state
  variable(s). \comment{could do with a figure here}

@@

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition?} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \rx reactor execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

The manner in which the agent interleaves planning \texttt{step}s and
synchronization for a single reactor is illustrated in
Fig.~\ref{fig:tick-exec}. Synchronization occurs at every tick while
interrupting planning; this allows a reactor to identify its current
state that can propagate through the reactor hierarchy in order to
ensure a consistent view of the state of the world between all the
reactors.
% This view of a reactor has two parallel tasks allowing for long
% planning deliberation while ensuring that the state evolution
% propagates through the agent as time advance.
These intertwined tasks together provide two natural information flows
(bottom-up for state estimation and top-down for plan projection) that
we expect to see in a control loop. As they are both inference-based
processes it is natural to implement a reactor based on generic
automated planning. For example in Fig. \ref{fig:agent}, apart from
\textsf{Shore comm.} and \textsf{Vehicle} acting as interfaces to the
system, every other reactor (in blue) are different instances of a \eu
reactor differentiated only by their model and \eu solver
configurations.

\subsection{The  \eu Deliberative Reactor}
\label{sec:arch:europa}

Within \rx, an instance of an \eu reactor comes with the sole purpose
to deliberate. A typical design of such a reactor takes its cues from
\texttt{IDEA} \cite{mus02, mus06} where both planning and execution
are based on a single unified model within a rich representation that
provides support for temporal inference. In \rx, the execution cycle
consists of two deliberation processes which are implemented as two
\eu solvers modifying the \emph{same} plan database. %  and for which the
% execution is managed by the \rx framework as time advance during the
% execution of the system:

@@

\item The \emph{synchronization solver} is a specialized \eu solver
  that integrates new information in the plan, the evolution of
  \emph{external} state variables and ensures that the reactor
  propagates these to identify the reactors current state as well as
  to inform other reactors of any state change on its \emph{internal}
  timelines. This solver is summoned at the beginning of every single
  tick.

\item The \emph{deliberation solver} manages the deliberation process
  of the reactor either to produce a new plan or alter its current
  plan as new goals are provided or when the synchronization solver
  identifies a conflict between current state and expectations from a
  previously generated plan. This process can span multiple ticks and
  can therefore be interrupted by the synchronization solver.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when the deliberation process eventually finds a
solution, synchronization is informed about generating the planned
state values \comment{should this read as ``...is informed about
  generated state values..'' instead?} on the {\em external} state
variables. This {\em external} plan defines the set of {\em goals} for
this state variable managed by another {\em reactor}.

Synchronization offers an important challenge for embedding planning
in a situated agent, typically %  as always been to go around the asumptions made by
% most classical planners. Specifically one
an assumption to restrict the problem to ``offline planning'' defined
in \cite{ghallab04} as follows:

\begin{quotation}
  The planner is not concerned with any change that may occur in
  $\Sigma$ \footnote{the world modeled by the plan domain} while it is
  planning; it plans for the given initial and goal states regardless
  of the current dynamics, if any. 
\end{quotation}

This assumption, while helpful to reduce the scope of the planning
problem, becomes problematic for planning within a highly dynamic
environment. In a typical robot this implies that the only point where
the planner can be informed about the current world state is prior to
the commencement of a planning cycle by giving to the planner the
initial state and current objectives. During the planning, it is
assumed that the agent can maintin the world as perceived by the
planner as stable. \cite{lemai04, lemai-chenevier2004} attempts to
reduce the impact of planning by allowing local plan repair in the
existing plan. In the eventuality that such repair is not feasible
they require that their robot comes to a full stop until (re-)planning
is complete. \comment{I may need more examples/refs here}. As Fig.
\ref{fig:tick-exec} shows, this assumption is problematic for our
architecture. Synchronization that tracks state evolution can occur
several times during the reactor deliberation process. By making
synchronization and deliberation share the same plan we provide a
solution that relaxes this static world assumption.

In the following sections we show how synchronization and deliberation
are implemented independently to show how their interaction through
the plan helps relax this assumption by making planning fully
integrated in the reactor execution cycle while potentially allowing
for more informed planning.
